### 2023-05-03 application.tex
up

\begin{proof}
Consider a given $k \in \N$.
Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
Consider a given Vector addition system, which has a finite number of transitions.
Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.

$\Bad$ admits a finite basis $B_\Bad$.
Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

@@

Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold.
\end{proof}

@@

\begin{proposition}
{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$ is decidable
\end{proposition}

\begin{proof}
Consider a given $k \in \N$.
Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
% $\Safe \subseteq Q \times \N^d$
$\Safe = S \setminus \Bad$
 downward-closed.
Consider a given Vector addition system with states, which has a finite number of transitions.
Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.

$\Bad$ admits a finite basis 
$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
q_{i_m}(\textbf{v}_{i_m}) \}$.
Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
For all $q \in Q$, all configurations reachable from $q(\textbf{u}_k)$ in $k$ or less steps are 
of the form $q'(\textbf{v}')$ with $ \textbf{v}' > \textbf{v}_{\Bad}$.

For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then $q'(\omega)$ is in the complement of $\Bad$, i.e. $\Safe$.

Let us assume from now on $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
all subsets of $\Safe$.
Because these subsets are upward-closed, we can compute a basis of
$\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
$\pred^{\leq k}(q_{j_\ell}(\omega))$.

We now consider the set $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
% \mathieu{Is it actually possible to check whether or not it is finite ?}
% If it is finite then w
We can check for the elements in it whether or not it is possible to reach from it $\Safe$ in $k$ or less steps.
If the set is finite then we stop the procedure once we have checked for every element.
If it is infinite then
$k$-resilience does not hold 
and there exists an element of the form $q(u)$
that is in $\Bad$ but not in any of the $\pred^{\leq k}(q_{j_i}(\omega))$
for which it is not possible to reach the `smaller' elements of $\Safe$ in $k$ or less steps and thus it is not possible at all to reach $\Safe$ in $k$ or less steps.
\end{proof}

\mathieu{À priori on devrais avoir la bounded resilience aussi.
En fait, comme on a strong monotony alors
 $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$
 converge vers
  $\Bad \setminus (\pred^{*}(q_{j_1}(\omega)) \cup \ldots \pred^{*}(q_{j_\ell}(\omega)) )$.
  Une fois que le set est stable alors, encore une fois, si il est infini on n'a pas la $k$-résilience
(pour tout $k$ à partir de ce rang là, du coup)
et si il est fini alors on peut vérifier, pour chaque élément, si il est possible d'atteindre
$\Safe$. Si il est possible pour chaque élément alors on prends le maximum des chemins depuis chacun de ces éléments et on fait le max de ça et le $k$ qui permet d'atteindre $\pred^{\leq k} = \pred^*$ pour les $q(\omega)$ ci-dessus et ça nous donne un $k_{\max}$ pour lequel on a la $k_{\max}$-résilience.
Si il est pas possible pour chaque élément alors ça veut dire il y a un élément qui atteindra jamais $\Safe$ et on n'a pas la bounded resilience ni la résilience d'ailleurs. 
Il faut vérifier que c'est possible pour une configuration de vérifier qu'elle est capable d'atteindre $\Safe$ downward closed. 
Normalement c'est le cas d'après Okzan (ils disent c'est équiv du submarking pb).}

>>>>

\begin{proof}
Consider a given $k \in \N$.
Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
Consider a given vector addition system $V$, which has a finite number of transitions.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.

The set $\Bad$ admits a finite basis $B_\Bad$.
Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

@@

Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
\end{proof}

@@

\begin{proposition}
{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
\end{proposition}

\begin{proof}
Consider a given $k \in \N$.
Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
% $\Safe \subseteq Q \times \N^d$
$\Safe = S \setminus \Bad$
 downward-closed.
Consider a given VASS $V$, which has a finite number of transitions.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.

The set $\Bad$ admits a finite basis 
$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
q_{i_m}(\textbf{v}_{i_m}) \}$.
Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
For all $q \in Q$, all configurations reachable from $q(\textbf{u}_k)$ in $k$ or less steps are 
of the form $q'(\textbf{v}')$ with $ \textbf{v}' > \textbf{v}_{\Bad}$.

For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
% $q'(\omega)$
$\uparrow q'(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$.

Let us assume from now on 
% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
 $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
all subsets of $\Safe$.
Because these subsets are upward-closed, we can compute a basis of
% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.

We now consider the set 
% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
% \mathieu{Is it actually possible to check whether or not it is finite ?}
% If it is finite then w
We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
If the set is finite then we stop the procedure once we have checked for every element.
If it is infinite then
$k$-resilience does not hold 
and we eventually find a witness that $k$-resilience does not hold.
Indeed, 
if
$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
is
infinite
then
there exists an element of the form $q(\textbf{u} )$
with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
and hence from which it is not possible at all to reach
$\Safe$ in $k$ or less steps.
\end{proof}

\iffalse
\mathieu{À priori on devrais avoir la bounded resilience aussi.
En fait, comme on a strong monotony alors
 $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$
 converge vers
  $\Bad \setminus (\pred^{*}(q_{j_1}(\omega)) \cup \ldots \pred^{*}(q_{j_\ell}(\omega)) )$.
  Une fois que le set est stable alors, encore une fois, si il est infini on n'a pas la $k$-résilience
(pour tout $k$ à partir de ce rang là, du coup)
et si il est fini alors on peut vérifier, pour chaque élément, si il est possible d'atteindre
$\Safe$. Si il est possible pour chaque élément alors on prends le maximum des chemins depuis chacun de ces éléments et on fait le max de ça et le $k$ qui permet d'atteindre $\pred^{\leq k} = \pred^*$ pour les $q(\omega)$ ci-dessus et ça nous donne un $k_{\max}$ pour lequel on a la $k_{\max}$-résilience.
Si il est pas possible pour chaque élément alors ça veut dire il y a un élément qui atteindra jamais $\Safe$ et on n'a pas la bounded resilience ni la résilience d'ailleurs. 
Il faut vérifier que c'est possible pour une configuration de vérifier qu'elle est capable d'atteindre $\Safe$ downward closed. 
Normalement c'est le cas d'après Okzan (ils disent c'est équiv du submarking pb).}
\fi

\begin{proposition}
{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
\end{proposition}

\begin{proof}
$\Bad$ admits a finite basis 
$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
q_{i_m}(\textbf{v}_{i_m}) \}$.
For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
% $q'(\omega)$
$\uparrow q'(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$.

Let us assume from now on 
% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
 $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
all subsets of $\Safe$.
Because these subsets are upward-closed, we can compute a basis of
% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.

We now consider the set 
% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.

There exists some $k_0 \in \N$ for which
and hence
$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.

From $k_0$ onwards ...
$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
is stationnary. 
If the set is finite then we stop the procedure once we have checked for every element
whether it is possible to reach $\Safe$ from it.
If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
hold
for $k_m = \max(k_0, k_{\pi})$ with 
$k_{\pi}$ the maximum of the length of the paths from 
$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
If it is infinite then
for every $k \in \N$,
$k$-resilience does not hold, 
and we eventually find a witness that $k$-resilience does not hold.
Indeed, 
if
$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
is
infinite
then
there exists an element of the form $q(\textbf{u} )$
with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
and hence from which it is not possible at all to reach
$\Safe$ in $k$ or less steps.
\end{proof}
